---
title: 'Final Project: Group 5. Effect of Russian Trolls on the 2016 U.S. Presidential election.'
author:
  - Ryan Appel
  - Angel Arribas
  - Drew Breyer
  - Humza Kahn
  - Devanshi Patel
  - Hanna Shin
  - Poonam Siyag
output:
  pdf_document: default
  html_document:
    df_print: paged
---
# Introduction
## Getting started

Loading packages is boring and time-consuming. First, you need to install packages. Second, you need to run them in R's environment.
```{r, include=FALSE}
if (!require("pacman")) install.packages("pacman")
p_load(data.table, dplyr) # two modern packages to modify data objects in R
p_load(stargazer) # a package to summaries regression results 
p_load(ggplot2, ggpubr) # two packages to visualize data
`%+%` <- function(x,y){paste0(x, y)} # collapse two strings into one
theme_set(theme_pubr(border = TRUE)) # make figures beautiful and ascetic by default
```

## Custom Functions

```{r}
# The below function provides a percentage after being passed a part and whole data.table
percentage <- function(part, whole, rounding.digits = 3) {
    part.count <- nrow(part)
    whole.count <- nrow(whole)
    percentage <- part.count / whole.count
    percentage <- percentage * 100
    percentage <- round(percentage, rounding.digits)
    return (percentage)
}

print.percentage <- function(part, whole, rounding.digits = 3) {
    p <- percentage(part, whole, rounding.digits)
    print( p %+% "%")
}
```

## Loading data
Load dataset `d` and inspect its structure.
```{r}
path <- 'https://raw.githubusercontent.com/RappelBerryPi/PSCI6303FinalProject/main/0-data/experiment_data.csv'
d <- fread(path)
```

let's start with summarizing the data
```{r}
glimpse(d)
percentage.control <- percentage(d[treated == FALSE,], d)
percentage.treatment <- percentage(d[treated == TRUE,], d)
percentage.control.changed <- percentage(d[treated == FALSE & voting_plan_2024 != general_voting_preference,], d[treated == FALSE,])
percentage.treatment.changed <- percentage(d[treated == TRUE & voting_plan_2024 != general_voting_preference,], d[treated == TRUE,])
```

now let's do a DID by looking at our variables: treatment and number_of_troll_tweets_seen vs if their 2024_voting_plan is the same as their general_voting_preference. First we have to create some extra columns.
```{r}
mean.troll.tweets <- mean(d$number_of_troll_tweets_seen)
d <- d[,voting_plan_different := voting_plan_2024 != general_voting_preference]
d <- d[,affected := number_of_troll_tweets_seen > mean.troll.tweets]
model.did <- lm(data = d, voting_plan_different ~ treated + affected + treated:affected)
stargazer(model.did, type = "text")
```

```{r}
print.percentage(d[affected == TRUE,],d)
print.percentage(d[affected == FALSE,],d)
print.percentage(d[treated == TRUE,],d)
print.percentage(d[treated == FALSE,],d)
```
